---
title: "Transformation"
author: "alex"
date: "5/22/2020"
output: html_document
---


```{r include = FALSE}
library(nycflights13)
flights <- nycflights13::flights
library(dplyr)

```

## Filter function  
Find all flights that:

Had an arrival delay of two or more hours
```{r}
filter(flights, arr_delay >= 120)
```

Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest %in% c("IAH", "HOU"))
```

Were operated by United (UA), American (AA), or Delta (DL)
```{r}
filter(flights, carrier %in% c("US", "AA", "DL"))
```


Departed in summer (July, August, and September)
```{r}
filter(flights, month %in% c(7,8,9))
```

Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, arr_delay > 120 & dep_delay <= 0)
```

Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60, dep_delay - arr_delay > 30)
```

Departed between midnight and 6am (inclusive)
```{r}
filter(flights, dep_time %% 2400 <= 600)
```
How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
```{r}
filter(flights, is.na(dep_time))
```




## Arrange Function  
How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(flights, desc(is.na(dep_time)), dep_time)
```


Sort flights to find the most delayed flights. Find the flights that left earliest.

```{r}
arrange(flights, desc(dep_delay))
```


Sort flights to find the fastest (highest speed) flights.
```{r}
fastest_flights <- mutate(flights, mph = distance / air_time * 60)
head(arrange(fastest_flights, desc(mph)))
```

Which flights travelled the farthest? 
```{r}
arrange(flights, desc(distance))
```

Which travelled the shortest?
```{r}
arrange(flights, distance)
```

## Select function
set a vector of col names as a variable and use it to select columns withteh one_of function.
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```
There are a number of helper functions you can use within select():

-starts_with("abc"): matches names that begin with “abc”.

-ends_with("xyz"): matches names that end with “xyz”.

-contains("ijk"): matches names that contain “ijk”.

-matches("(.)\\1"): selects variables that match a regular expression. This one matches any variables that contain repeated characters. You’ll learn more about regular expressions in strings.

-num_range("x", 1:3): matches x1, x2 and x3.

## mutate function
If you only want to keep the new variables, use transmute():
```{r}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

Integer division operator %/%
```{r}
1504 %/% 100
```
Modulo operator %%
```{r}
1504 %% 100
```
Combine the two to get total minutes ofter midnight (1440 = 24 hours)
```{r}
time2mins <- function(x) {
  (x %/% 100 * 60 + x %% 100) %% 1440
}
flights_times <- mutate(flights,
  dep_time_mins = time2mins(dep_time),
  sched_dep_time_mins = time2mins(sched_dep_time))
```









