---
title: "beer"
author: "Alex MacCalman"
date: "8/29/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Shows how to use bootstraping to understand relationships for inference (not predicting).
Our modeling goal here is to estimate how much sugar beer producers use relative to malt.  
```{r}
library(tidyverse)
#import the data
brewing_materials_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-31/brewing_materials.csv")
# let's look at teh types of materials
brewing_materials_raw %>%  count(type, sort = TRUE)

#This looks at the amount of barrels (month_count) used for each material
brewing_materials_raw %>% count(type, wt = month_current, sort = TRUE) # wt parameter sums each group type

```
how much sugar do beer producers need for every barrel of malt? Estiamte the amount of sugar beer producers use per barrel?  
```{r}
# explore, filter a subset of materials, take out bad data (year >=2016, and December)
# try using a ploty to identify the bad data and filter it out that way.
brewing_filtered <- brewing_materials_raw %>% 
        filter(type %in% c("Malt and malt products",
                           "Sugar and syrups",
                           "Hops (dry)"),
               year < 2016,
               month != 12) %>% 
        mutate(date = paste0(year, "-", month, "-01"),
               date = lubridate::ymd(date))

brewing_filtered %>% 
        ggplot(aes(date, month_current, color = type)) +
        geom_point()

#reshape the data and clean the col names
brewing_materials <- brewing_filtered %>% 
        select(date, type, month_current) %>% 
        pivot_wider(names_from = type, values_from = month_current) %>% 
        janitor::clean_names()

# look at the relationship between malt and sugar
brewing_materials %>% 
        ggplot(aes(malt_and_malt_products, sugar_and_syrups)) +
        geom_point() +
        geom_smooth(method = "lm")

#now we will fit a simple linear model
beer_fit <- lm(sugar_and_syrups ~ 0 + malt_and_malt_products,
               data = brewing_materials)

library(tidymodels)
tidy(beer_fit)
```
# looks like a linear regression model would not hold up to this type of data.  
## Bootstrap resampling
```{r}
set.seed(123)
#create 1000 bootstraps, plus one that is the whole (apparent)
beer_boot <- bootstraps(brewing_materials,
           times = 1e3,
           apparent = TRUE)
# map apply the lm to all the bootstrap samples by adding a column
beer_models <- beer_boot %>% 
        mutate(model = map(splits, ~ lm(sugar_and_syrups ~ 0 + malt_and_malt_products,
               data = .)),
               coef_info = map(model, tidy))
# we have to unnest the coef_info column to break out the nested tibbles in beer_models
beer_coefs <- beer_models %>% 
        unnest(coef_info)
# this allows us to treat the results of the model as data to do inference on.
```
## Evaluate results  
```{r}
# create a histogram of the estiamts from the bootstrap samples
beer_coefs %>% 
        ggplot(aes(estimate)) +
        geom_histogram(apha = 0.7)
#now let's get the bootstrap confidence intervals
int_pctl(beer_models, coef_info) # to do this we needed to set the apparent to TRUE when creating the bootstraps. the coef_info is the statistics parameter from the beer_models tibble

```
We can also visualize some of these fits to the bootstrap resamples. First, let’s use augment() to get the fitted values for each resampled data point. 
make a new column. broom has tidy, glance, and augment. Augment goes back to the original data points and add things to them.
```{r}
beer_aug <- beer_models %>%
  sample_n(200) %>% #only get 200
  mutate(augmented = map(model, augment)) %>% # adds columns to a dataset, containing information such as fitted values, residuals or cluster assignments. All columns added to a dataset have . prefix to prevent existing columns from being overwritten.
  unnest(augmented) # this statement unnests the tibble that was augmented.

#now let's visualize
ggplot(beer_aug, aes(malt_and_malt_products, sugar_and_syrups)) +
  geom_line(aes(y = .fitted, group = id), alpha = .2, col = "cyan3") +
  geom_point()
```

